### 2024-11-04 README.Rmd
package files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# canaccessR

<!-- badges: start -->
<!-- badges: end -->

Package {canaccessR} is an [Open Data Product](https://rdcu.be/dY1Ic) with travel time tables to employment and groceries in selected Canadian Metropolitan Areas.

The package was initially developed for a writing project commissioned by the [School of Cities](https://schoolofcities.utoronto.ca) of the University of Toronto, and can be used to replicate the results reported in the paper.

## Installation

You can install the development version of canaccessR like so:

``` r
# install.packages("remotes")
remotes::install_github("paezha/canaccessR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r load-packages}
library(canaccessR)
library(ggplot2)
library(sf)
```

Load the travel time table to employment for Halifax:
```{r load-data}
data(travel_matrix_emp_hal)
```

The median travel time in minutes from all origins to all destinations in Halifax is:
```{r}
median(travel_matrix_emp_hal$travel_time_p50)
```

Simple features objects for cartography are also included. For example, this is Quebec:
```{r}
data("census_data_da_que")
```

Plot the population by Dissemination Area:
```{r}
ggplot() +
  geom_sf(data = census_data_da_que,
          aes(fill = Population)) +
  scale_fill_fermenter(palette = "Reds",
                       direction = 1) +
  theme_minimal()
```