### 2024-11-04 data-raw/00_census_data.qmd

---
title: "00_census_data"
format: html
---

```{r}
#| label: load packages
library(cancensus) # Access, Retrieve, and Work with Canadian Census Data and Geography
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(furrr) # Apply Mapping Functions in Parallel using Futures
library(fs) # Cross-Platform File System Operations Based on 'libuv'
library(sf) # Simple Features for R
library(tidyverse) # Easily Install and Load the 'Tidyverse'

cache_path <- fs::dir_create("./cache")
data_path <- fs::dir_create("./data")
data_raw_path <- fs::dir_create("./data-raw")

# See here for instructions to obtain an API key: https://mountainmath.github.io/cancensus/index.html#api-key
options(cancensus.api_key = "YOUR_KEY")
options(cancensus.cache_path = cache_path)
```

# Census Data
## Define CMA lists for iterative getting of data

```{r}
# list of CMA IDs for cancensus
cma_geouid_list <- c(
  "Toronto" = "35535", 
  "Montréal" = "24462", 
  "Vancouver" = "59933",
  "Calgary" = "48825",
  "Ottawa" = "505",
  "Edmonton" = "48835",
  "Québec City" = "24421",
  "Winnipeg" = "46602",
  "Hamilton" = "35537",
  "Waterloo" = "35541",
  "London" = "35555",
  #"Niagara" = "35539",
  "Halifax" = "12205",
  "Oshawa" = "35532",
  "Abbotsford" = "59932")

# create list of analysis regions - collapsing some into others e.g. oshawa into toronto
study_region_names_list <- c(
  "Toronto" = "35535", 
  "Montréal" = "24462", 
  "Vancouver" = "59933",
  "Calgary" = "48825",
  "Ottawa" = "505",
  "Edmonton" = "48835",
  "Québec City" = "24421",
  "Winnipeg" = "46602",
  "Hamilton" = "35537",
  "Waterloo" = "35541",
  "London" = "35555",
  #"Niagara" = "35539",
  "Halifax" = "12205",
  "Toronto" = "35532",
  "Vancouver" = "59932") |> 
  enframe() |> 
  select(2:1) |> 
  deframe()

# ordered factor of study regions for figures and mapping
study_region_names_order_list <- list(
  "Toronto" = "gta", 
  "Montréal" = "mtl", 
  "Vancouver" = "van",
  "Calgary" = "cal",
  "Ottawa" = "ott",
  "Edmonton" = "edm",
  "Québec City" = "que",
  "Winnipeg" = "win",
  "Hamilton" = "ham",
  "Waterloo" = "wat",
  "London" = "ldn",
  "Halifax" = "hal") |> 
  enframe() |> 
  select(2:1) |> 
  deframe()

# define analysis regions - use to collapse CMAs into larger analysis regions
r5r_analysis_region_list <- c(
  "ggh" = "35535", 
  "mtl" = "24462", 
  "van" = "59933",
  "cal" = "48825",
  "ott" = "505",
  "edm" = "48835",
  "que" = "24421",
  "win" = "46602",
  "ggh" = "35537",
  "ggh" = "35541",
  "ldn" = "35555",
  "ggh" = "35539",
  "hal" = "12205",
  "ggh" = "35532",
  "van" = "59932") |> 
  enframe() |> 
  select(2:1) |> 
  deframe()

r5r_named_networks <- list(
  "ggh" = "ggh", 
  "mtl" = "mtl", 
  "van" = "van",
  "cal" = "cal",
  "ott" = "ott",
  "edm" = "edm",
  "que" = "que",
  "win" = "win",
  "ldn" = "ldn",
  "hal" = "hal")
```

## Define MJ Census Vectors for 2016 Census

```{r}
mj_vectors_2016 <- c(
  "tot_age_popn" = "v_CA16_1",
  "age_0_to_14" = "v_CA16_4",
  "age_15_to_64" = "v_CA16_61",
  "age_65_plus" = "v_CA16_244",
  "tot_cens_fam_priv_hh" = "v_CA16_484",
  "fem_parent" = "v_CA16_489",
  "LICO_AT" = "v_CA16_2555", 
  "tot_LICO_status_2015" = "v_CA16_2510",
  "knwl_off_lang" = "v_CA16_512",
  "neither_EN_FR" = "v_CA16_524",
  "Aborin_ID_priv_HH" = "v_CA16_3852",
  "Aborin_ID" = "v_CA16_3855",
  "tot_priv_hh_hsing_suit" = "v_CA16_4859",
  "not_suitable" = "v_CA16_4861",
  "dwell_condtns" = "v_CA16_4870",
  "maj_rep_needed" = "v_CA16_4872",
  "tot_shelt_cost" = "v_CA16_4886",
  "shelt_gt30pc" = "v_CA16_4888",
  "tot_tenant_HH" = "v_CA16_4897",
  "tenant_HH_subsid" = "v_CA16_4898",
  "tot_immig_stat" = "v_CA16_3405",
  "immig_stat_last5" = "v_CA16_3432",
  "tot_adm_immig_popn_1980_to_2016" = "v_CA16_3831",
  "refugees" = "v_CA16_3846",
  "visible_min" = "v_CA16_3957",
  "tot_visible_min" = "v_CA16_3954",
  "Black" = "v_CA16_3966",
  "tot_lbr_f_15_plus" = "v_CA16_5597",
  "lbr_f_unemp" = "v_CA16_5606",
  "tot_commuting_dur_lbr_f_15_plus" = "v_CA16_5813",
  "commuting_dur_60mins_plus" = "v_CA16_5828",
  "ave_HH_size" = "v_CA16_425",
  "tot_degree_age_15_plus" = "v_CA16_5051",
  "bachelor_or_higher" = "v_CA16_5078",
  "mhh_inc" = "v_CA16_2397",
  "popn_dens" = "v_CA16_406",
  "from_mode_total" = "v_CA16_5792",
  "from_mode_car" = "v_CA16_5795",
  "from_mode_pass" = "v_CA16_5798",
  "from_mode_tran" = "v_CA16_5801",
  "from_mode_walk" = "v_CA16_5804")
```

## Get CMA Data and Geometries

```{r}
census_data_cma <- map_dfr(
  cma_geouid_list, 
  ~ get_census(
  dataset = 'CA16', 
  regions = list(CMA = c(.x)),
  vectors = mj_vectors_2016,
  level = 'CMA',
  geo_format = "sf",
  use_cache = TRUE)) |> 
  mutate(
    r5r_analysis_region = as.factor(r5r_analysis_region_list[GeoUID]),
    study_region_name = as.factor(study_region_names_list[GeoUID]),
    study_region_name = fct_relevel(study_region_name, study_region_names_order_list)
  )
```

join water erased geometries

```{r}
cmas_no_water <- st_read(glue::glue(here::here(), 
                                    "/data-raw/census_data_cma_erased.shp"))

census_data_cma <- census_data_cma |> 
  st_drop_geometry() |> 
  left_join(cmas_no_water, by = "GeoUID") |> 
  st_as_sf() |> 
  st_make_valid() |> 
  st_transform(crs = 4326) 
```

## Get DA Data and Geometries

```{r}
census_data_da <- map_dfr(
  cma_geouid_list, 
  ~ get_census(
  dataset = 'CA16', 
  regions = list(CMA = c(.x)),
  vectors = mj_vectors_2016,
  level = 'DA',
  geo_format = "sf",
  use_cache = TRUE)) |> 
  mutate(
    study_region_name = study_region_names_list[CMA_UID],
    study_region_name = fct_relevel(study_region_name, study_region_names_order_list),
    r5r_analysis_region = r5r_analysis_region_list[CMA_UID]
    )
```

```{r}
das_no_water <- st_read(glue::glue(here::here(), 
                                    "/data-raw/census_data_da_erased.shp"))

# DA_EMP_2016_Canada.csv is a confidential file
da_employment <- read_csv(glue::glue(here::here(), 
                                    "/data/DA_EMP_2016_Canada.csv")) |> 
  select(GeoUID, Total_Emp) |> 
  mutate(GeoUID = as.character(GeoUID))

census_data_da <- census_data_da |> 
  left_join(da_employment |> select(GeoUID, Total_Emp), by = "GeoUID") |> 
  mutate(Total_Emp = replace_na(Total_Emp, 0)) |> 
  st_drop_geometry() |> 
  left_join(das_no_water, by = "GeoUID") |>
  st_as_sf() |> 
  st_make_valid() |> 
  st_transform(crs = 4326)
```

## Region Backgrounds for Mapping

```{r}
region_backgrounds <- st_read(glue::glue(here::here(), 
                                    "/data-raw/pr_region_backgrounds.shp")) |> 
  rename(region_name = regn_nm)
```

## Save Census Data

```{r}
# save(
#   census_data_da, 
#   , 
#   region_backgrounds, 
#   file = "./data/census_data.Rdata")
```

Regions used for routing in {r5r}:
```{r}
regions <- levels(census_data_cma$r5r_analysis_region)
```

Save CMA data by region:
```{r}

for(i in regions){
  census_data_cma |>
    rename(region = r5r_analysis_region) |>
    filter(region == i) |>
    select(Type, GeoUID, name, region) |>
    assign(x = glue::glue("census_data_cma_{i}"),
         value = _)

  save(list = glue::glue("census_data_cma_{i}"),
       file = glue::glue(here::here(),
                       "/data/census_data_cma_{i}.rda"),
     compress = "xz")
}

```

Save DA data by region:
```{r}
for(i in regions){
  census_data_da |>
    rename(region = r5r_analysis_region) |>
    filter(region == i) |>
    select(-c(`Shape Area`, `Region Name`, `Area (sq km)`, popn_dens, `study_region_name`)) |>
    select(Type, GeoUID, region, everything()) |>
    assign(x = glue::glue("census_data_da_{i}"),
         value = _)

  save(list = glue::glue("census_data_da_{i}"),
       file = glue::glue(here::here(),
                       "/data/census_data_da_{i}.rda"),
     compress = "xz")
}

```

Save backgrounds by region:
```{r}
for(i in regions){
  region_backgrounds |>
    rename(region = regn_xp) |>
    select(region) |>
    filter(region == i) |>
    assign(x = glue::glue("region_background_{i}"),
         value = _)

  save(list = glue::glue("region_background_{i}"),
       file = glue::glue(here::here(),
                       "/data/region_background_{i}.rda"),
     compress = "xz")
}

```

Save all regions in a single object. These files are pretty heavy, so we will put them in a confidential folder:
```{r}
save(
  census_data_da,
  file = glue::glue(here::here(), "/data-raw/census/census_data_da.rda"),
  compress = "xz")
```

```{r}
save(
  census_data_cma, 
  file = glue::glue(here::here(), "/data-raw/census/census_data_cma.rda"),
  compress = "xz")
```

```{r}
save(
  region_backgrounds, 
  file = glue::glue(here::here(), "/data-raw/census/region_backgrounds.rda"),
  compress = "xz")
```