### 2024-11-04 data-raw/04_accessibility_measure_calcs.qmd

---
title: "05_accessibility_measure_calcs"
format: html
---

# Setup

```{r}
#| label: load packages

#library(arrow) # Integration to 'Apache' 'Arrow'
library(duckdb) # DBI Package for the DuckDB Database Management System
library(dbplyr) # A 'dplyr' Back End for Databases
library(fs) # Cross-Platform File System Operations Based on 'libuv'
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(hms) # Pretty Time of Day
library(sf) # Simple Features for R
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tmap) # Thematic Maps
#library(accessibility) 
```

# Load Data

```{r}
# Load census data. These files are confidential and were created in notebook 00_census_data.qmd
load(file = glue::glue(here::here(), 
                       "/data-raw/census/census_data_da.rda"))
load(file = glue::glue(here::here(), 
                       "/data-raw/census/census_data_cma.rda"))
load(file = glue::glue(here::here(), 
                       "/data-raw/census/region_backgrounds.rda"))

census_data_da <- census_data_da |>  
  mutate(Population = ifelse(is.na(Population), 0, Population),
          Total_Emp = ifelse(is.na(Total_Emp), 0, Total_Emp))

# Load grocers data. This is a confidential file that was created in notebook 01_cleaning_destinations.qmd
load(glue::glue(here::here(),
                       "/data-raw/DMTI_2023_EPOI/grocers.rda"))
```

We will only use the 12pm departure transit travel time matrices (already run in 03.qmd) to grocery stores (DA-to-store) and employment locations (DA-to-DAs). This will be the basis of our accessibility calculations.

Census data by DA for join
```{r}
census_data_da_for_join <- census_data_da |> 
  st_drop_geometry() |> 
  select(GeoUID, study_region_name, Population) #|>  #including POPULATION here
  #arrow::to_duckdb()
```

# Travel time to grocery stores

First, grocery store ttms:

<!--

**WARNING**: The travel time matrices for 12 pm are missing.

```{r}
#| eval: false
ttm_small_path <- fs::dir_create("./ttms_12pm")

ttm_12pm_folders <- fs::dir_ls("./ttms", 
                               type = "directory", 
                               regexp = "12_00_00", 
                               recurse = TRUE)

walk(ttm_12pm_folders, ~dir_copy(.x, fs::path(ttm_small_path, (.x)), overwrite = TRUE))
```

read in travel time matrices as a data frame:
```{r}
# preparing the data frame to only consider only the 12:00pm departure (+15 min departure window), on Saturday!
travel_matrix_grocer <- arrow::open_dataset(glue::glue(here::here(), 
                                                       "/data-raw/ttms_12pm")) |>
  collect() |>
  mutate(
    start_datetime = lubridate::ymd_hms(start_datetime)
    ) |> filter(start_datetime  == as.POSIXct("2019-04-20 12:00:00", tz="UTC") |  
         start_datetime  == as.POSIXct("2019-06-29 12:00:00", tz="UTC") |  #for quebec
           start_datetime  == as.POSIXct("2023-04-22 12:00:00", tz="UTC")) |> #keeping saturdays!

  mutate(start_year = lubridate::year(start_datetime)) |> #adding the start_year

  left_join(census_data_da_for_join |> select(c(GeoUID, study_region_name, Population)), by = c("from_id" = "GeoUID")) #joining the population to the "from_id"

```

Assigning grocery store related travel impedance weights, for now, from Kwan 1998
```{r}
b0 <- 180
travel_matrix_grocer <- travel_matrix_grocer |>  
  mutate(
    o_j = 1, # opportunity intensity of each grocery store is 1
    f = exp(-travel_time_p50 ^ 2 / b0) #informed by Kwan 1998) 
    )
```

## grocer spatial accessibility

Let's also add non-competitive grocery store spatial accessibility to the frame:
```{r}
access_grocers_2019 <- travel_matrix_grocer_2019 |> group_by(from_id) |> summarise(S_i = sum(f*o_j))
access_grocers_2023 <- travel_matrix_grocer_2023 |> group_by(from_id) |> summarise(S_i = sum(f*o_j))

#also adding per capita spatial accessibility 
spatial_mea_grocers_2019_DA <- spatial_mea_grocers_2019_DA |> left_join(access_grocers_2019, by="from_id") |>
  mutate(s_i = S_i/Population)
spatial_mea_grocers_2023_DA <- spatial_mea_grocers_2023_DA |> left_join(access_grocers_2023, by="from_id") |>
  mutate(s_i = S_i/Population)
```
```{r}
#to save space, remove ttms grocers
rm(travel_matrix_grocer, travel_matrix_grocer_2019, travel_matrix_grocer_2023)
```
-->

# Accessibility to employment

Read DA-to-DA travel time matrices as a dataframe:
```{r}
# preparing the data frame to only consider only the 12:00pm departure (+15 min departure window), on Tuesday!
travel_matrix_employ <- arrow::open_dataset(glue::glue(here::here(), 
                                                       "/data-raw/ttms_employment")) |>
  collect() |>
  mutate(
    start_datetime = lubridate::ymd_hms(start_datetime)
    ) |>
  filter(start_datetime  == as.POSIXct("2019-04-16 08:00:00", tz="UTC") |  
           start_datetime  == as.POSIXct("2019-06-25 08:00:00", tz="UTC") |  #quebec!
           start_datetime  == as.POSIXct("2023-04-18 08:00:00", tz="UTC")) |> #keeping tuesdays!

  mutate(start_year = lubridate::year(start_datetime)) |> #adding the start_year

  left_join(census_data_da_for_join |> select(c(GeoUID, study_region_name, Population)), by = c("from_id" = "GeoUID")) |> #joining the population to the "from_id"

  left_join(census_data_da |> st_drop_geometry() |> select(c(GeoUID, Total_Emp)), by = c("to_id" = "GeoUID")) #joining the JOBS to the "to_id"

```

Save travel time matrices for employment:
```{r}
regions <- levels(as.factor(travel_matrix_employ$region))

for(i in regions){
  travel_matrix_employ |>
    filter(region == i) |>
    assign(x = glue::glue("travel_matrix_emp_{i}"),
         value = _)

  save(list = glue::glue("travel_matrix_emp_{i}"),
       file = glue::glue(here::here(),
                       "/data/travel_matrix_emp_{i}.rda"),
     compress = "xz")
}

```

<!--  
Assigning employment related travel impedance weights, for now, for transit mode in Kapitllian et al. 2023 is used:

```{r}
travel_matrix_employ <- travel_matrix_employ |>  mutate(
    f = case_when(
      study_region_name == "Calgary" ~  (1 / (1 + (travel_time_p50 / 41.6) ^ 5.0643)),
      study_region_name == "Edmonton" ~  (1 / (1 + (travel_time_p50 / 53) ^ 5.0643)),
      study_region_name == "Toronto" ~ (1 / (1 + (travel_time_p50 / 49) ^ 4.4856)),
      study_region_name == "Halifax" ~  (1 / (1 + (travel_time_p50 / 49) ^ 4.2019)),
      study_region_name == "London" ~ (1 / (1 + (travel_time_p50 / 47) ^ 4.6801)),
      study_region_name == "Montréal" ~ (1 / (1 + (travel_time_p50 / 44) ^ 4.1168)),
      study_region_name == "Ottawa" ~  (1 / (1 + (travel_time_p50 / 42.2) ^ 5.0643)),
      study_region_name == "Québec City" ~ (1 / (1 + (travel_time_p50 / 49) ^ 4.1947)),
      study_region_name == "Vancouver" ~ (1 / (1 + (travel_time_p50 / 45) ^ 4.1339)),
      study_region_name == "Winnipeg" ~ (1 / (1 + (travel_time_p50 / 45) ^ 4.6869)),
      study_region_name == "Hamilton" ~ (1 / (1 + (travel_time_p50 / 50) ^ 4.6869)),
      study_region_name == "Waterloo" ~ (1 / (1 + (travel_time_p50 / 40.3) ^ 4.6801)),
      .default = 9999
    ))
```

## job spatial availability

Calculating employment spatial_availability for 2019:
```{r}
#for 2019
travel_matrix_employ_2019 <- travel_matrix_employ |> filter(start_year == "2019")

spatial_mea_employ_2019 <- 
  travel_matrix_employ_2019 |> mutate(catch = 1) |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = Population,
                    opp = Total_Emp,
                    r = catch, #value of 1, all are in catchment.
                    f = f, 
                    alpha = 1) |>
  group_by(from_id) |>
  summarize(V_i = sum(V_ij, na.rm = TRUE))

#Checks, to make sure the total number of opportunities equals the total sum of V_i:
travel_matrix_employ_2019 |> group_by(to_id) |> summarise(jobs = first(Total_Emp)) |> select(jobs) |> sum()
spatial_mea_employ_2019$V_i |> sum()
```

for 2023:
```{r}
travel_matrix_employ_2023 <- travel_matrix_employ |> filter(start_year == "2023")

spatial_mea_employ_2023 <- 
  travel_matrix_employ_2023 |> mutate(catch = 1) |> 
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = Population,
                    opp = Total_Emp,
                    r = catch, #value of 1, all are in catchment.
                    f = f, 
                    alpha = 1) |>
  group_by(from_id) |>
  summarize(V_i = sum(V_ij, na.rm = TRUE))

#Checks, to make sure the total number of opportunities equals the total sum of V_i:
travel_matrix_employ_2023 |> group_by(to_id) |> summarise(jobs = first(Total_Emp)) |> select(jobs) |> sum()
spatial_mea_employ_2023$V_i |> sum()
spatial_mea_employ_2023|> filter(V_i == 0)

#NOTE: the total number of jobs (and population) differs between 2019 and 2023 -- this is because people/jobs with OD travel times are different from the 2019 vs. 2023 travel matrices. 
```

job spatial availability per capita:
```{r}
spatial_mea_employ_2019_DA <- spatial_mea_employ_2019 |> full_join(census_data_da, by=c("from_id" = "GeoUID")) |>
  mutate(Population = ifelse(is.na(Population), 0, Population),
         v_i = V_i/Population)

spatial_mea_employ_2023_DA <- spatial_mea_employ_2023 |> full_join(census_data_da, by=c("from_id" = "GeoUID"))|>
  mutate(Population = ifelse(is.na(Population), 0, Population),
         v_i = V_i/Population)
```

## job spatial accessibility

Let's add non-competitive job spatial accessibility to the frame:
```{r}
#calculating Hansen-type spatial accessibility, sum of (f*opportunities) for each origin (from_id) to all destination (possible to_ids):
access_employ_2019 <- travel_matrix_employ_2019 |> group_by(from_id) |> summarise(S_i = sum(f*Total_Emp))
access_employ_2023 <- travel_matrix_employ_2023 |> group_by(from_id) |> summarise(S_i = sum(f*Total_Emp))

#inspect:
access_employ_2019$S_i |> summary()
access_employ_2023$S_i |> summary()

#also adding per capita spatial accessibility 
spatial_mea_employ_2019_DA <- spatial_mea_employ_2019_DA |> left_join(access_employ_2019, by="from_id") |>
  mutate(s_i = S_i/Population)
spatial_mea_employ_2023_DA <- spatial_mea_employ_2023_DA |> left_join(access_employ_2023, by="from_id") |>
  mutate(s_i = S_i/Population)
```

```{r}
#to save space, remove ttms grocers
rm(travel_matrix_employ, travel_matrix_employ_2019, travel_matrix_employ_2023)
```

Saving all files
```{r, eval=FALSE}
#to save space.. remove the census data columns
spatial_mea_employ_2019_DA <- spatial_mea_employ_2019_DA |> select(c("from_id", "study_region_name", "V_i", "v_i", "S_i", "s_i"))
spatial_mea_employ_2023_DA <- spatial_mea_employ_2023_DA |> select(c("from_id", "study_region_name", "V_i", "v_i", "S_i", "s_i"))
spatial_mea_grocers_2019_DA <- spatial_mea_grocers_2019_DA |> select(c("from_id", "study_region_name", "V_i", "v_i", "S_i", "s_i"))
spatial_mea_grocers_2023_DA <- spatial_mea_grocers_2023_DA |> select(c("from_id", "study_region_name", "V_i", "v_i", "S_i", "s_i"))

save(spatial_mea_employ_2019_DA, 
     spatial_mea_employ_2023_DA , 
     spatial_mea_grocers_2019_DA, 
     spatial_mea_grocers_2023_DA , 
     file="./data/spatial_access_measures.Rda")
```