### 2024-11-04 data-raw/05_transit_statistics.qmd
Added table with population statistics

---
title: "transit statistics"
format: html
---

Libraries:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
library(lubridate) # Make Dealing with Dates a Little Easier
library(readr) # Read Rectangular Text Data
library(tidyr) # Tidy Messy Data
#library(sf) # Simple Features for R
#library(tibble) # Simple Data Frames
#library(tmap) # Thematic Maps
```

- Downloaded "Urban Public Transit, February 2024" table: https://www150.statcan.gc.ca/n1/daily-quotidien/240419/dq240419b-eng.htm and https://doi.org/10.25318/2310025101-eng

#Reading in and filtering

Reading in files
```{r}
transit_statistics <- read_csv(glue::glue(here::here(),
                                          "/data-raw/2310025101_databaseLoadingData.csv"))
```

```{r}
summary(transit_statistics)
```

Select the relevant information:
```{r}
transit_statistics <- transit_statistics |>
  transmute(GEO = factor(GEO),
            REF_DATE = lubridate::ym(REF_DATE) |>
              as.Date(),
         # Values are in millions
         VALUE = VALUE * 1000000,
         VAR = case_when(`Total revenue and total passenger trips` == "Total revenue, excluding subsidies" ~ "Revenue",
                         `Total revenue and total passenger trips` == "Total passenger trips" ~ "Ridership"))
```

Pivot to a wide table to have a column for revenue and other for ridership:
```{r}
transit_statistics <- transit_statistics |>
  pivot_wider(names_from = "VAR", values_from = "VALUE")
```

Save:
```{r}
save(transit_statistics,
     file = glue::glue(here::here(),
                       "/data/transit_statistics.rda"),
     compress = "xz")
```