### 2024-11-04 data-raw/06_population_statistics.qmd

---
title: "transit statistics"
format: html
---

Libraries:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(glue) # Interpreted String Literals
library(here) # A Simpler Way to Find Your Files
#library(lubridate) # Make Dealing with Dates a Little Easier
library(readr) # Read Rectangular Text Data
library(tidyr) # Tidy Messy Data
```

- Downloaded "Urban Public Transit, February 2024" table: https://www150.statcan.gc.ca/n1/daily-quotidien/240419/dq240419b-eng.htm and https://doi.org/10.25318/2310025101-eng

#Reading in and filtering

Reading in files
```{r}
population_statistics <- read_csv(glue::glue(here::here(),
                                          "/data-raw/population_statistics.csv"))
```

```{r}
summary(population_statistics)
```

Select the relevant information:
```{r}
population_statistics <- population_statistics |>
  transmute(size_rank = factor(`Size Rank`),
            year = 2021,
            cma_name = factor(`CMA Geographic Name`),
            province = factor(Province),
            population = `Population (2021)`)
```

Save:
```{r}
save(population_statistics,
     file = glue::glue(here::here(),
                       "/data/population_statistics.rda"),
     compress = "xz")
```